---
title: "Data 605 - Assginment 3"
author: "Amit Kapoor"
date: "2/14/2020"
output:
  pdf_document: default
  html_document: default
---

# Problem set 1

## 1. What is the rank of the matrix A?

$A = \left( \begin{matrix} 1&2&3&4 \\ -1&0&1&3 \\ 0&1&-2&1 \\ 5&4&-2&-3 \end{matrix} \right)$

```{r matix-A, echo=TRUE}
A <- matrix(cbind(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), byrow = T, ncol = 4)
A
```

```{r rank-A, echo=TRUE}

# to calc matrix rank using rankMatrix function from Matrix pkg
require(Matrix)

rankMatrix(A)[1]
```



## 2. Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

For mxn matrix where m>n, the rank would be n if all n column are linearly independent. Assuming matrix is non-zero, the minimum rank would be 1 given at least 1 pivot.

## 3. What is the rank of matrix B?

$B = \left( \begin{matrix} 1&2&1 \\ 3&6&3 \\ 2&4&2 \end{matrix} \right)$

```{r matix-B, echo=TRUE}
B <- matrix(cbind(1,2,1,3,6,3,2,4,2), byrow = T, ncol = 3)
B
```

Seeing the matrix B, row 2 is 3 times row 1 and row 3 is 2 times row 1 so we ended up having only one linear independent row therfore rank of B is 1. 

here is using R.

```{r rank-B, echo=TRUE}

rankMatrix(B)[1]
```



# Problem set 2

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

$A = \left( \begin{matrix} 1&2&3 \\ 0&4&5 \\ 0&0&6 \end{matrix} \right)$

Please show your work using an R-markdown document.

```{r define-A, echo=TRUE}
library(pracma)
A <- matrix(cbind(1,2,3,0,4,5,0,0,6), byrow = T, ncol = 3)
A
```


```{r charpoly-A, echo=TRUE}

# to computes the characteristic polynomial
charpoly(A)
```


```{r eigen-values-A, echo=TRUE}

eig <- eigen(A)

# eigen values of A
eig$values
```

```{r eigen-vectors-A, echo=TRUE}
# eigen vectors of A
eig$vectors
```




