---
title: "Data607 - Homework15"
author: "Amit Kapoor"
date: "5/12/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```


#### 1. Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.
**( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )**


```{r reg-model, echo=TRUE}

x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

reg_line <- lm(y~x)
reg_line
```


Per regression model, the line is **y =  -14.8 + 4.26x**

Below is the line plot.

```{r plot-line, echo=TRUE}

reg_line %>% 
  ggplot(aes(x, y)) +
  geom_point() +
  geom_smooth(method = lm, se = F)
```






#### 2. Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.
$$f(x, y) = 24x - 6xy^2 - 8y^3$$


**Partial derivatives:**

$$f_x(x, y) = 24 - 6y^2$$

$$f_y(x, y) = -12xy - 24y^2$$

**For Critical Points:**

$f_x = 0$ 

If $24-6y^2=0$, then $y^2 = 4$ and $y = \pm2$

**Substitute y values in $f_y=0$:**

If $y=2$ and $-12xy - 24y^2=0$, then $-24x = 24 * 4$ and $x=-4$

If $y=-2$ and $-12xy - 24y^2=0$, then $24x = 24 * 4$ and $x=4$

Calculate $f(x, y)$.

$$f(4,-2) = 24\times 4 - 6\times4\times (-2)^2 - 8 \times (-2)^3 = 64$$

$$f(-4,2) = 24\times (-4) - 6\times(-4)\times 2^2 - 8 \times 2^3 = -64$$

**Two critical points**: $(4,-2,64)$ and $(-4, 2, -64)$.

Now Lets use second derivative test to determine if points are minimum, maximum or saddle. 

**Second partial derivatives:**

$$f_{xx}=0$$

$$f_{yy}=-12x-48y$$

$$f_{xy}=-12y$$

Not Putting the values in formula 

$$D(x,y) = f_{xx} f_{yy}-f^2_{xy} = -(-12y)^2 = -144y^2$$

If we replace y=2 or y=-2, we get D<0 hence critical points (-4,2,-64) and (4,-2,64) are both saddle points.





#### 3. A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the "house" brand and 40 + 11x - 23y units of the "name" brand. 

* **Step 1. Find the revenue function R ( x, y )**

Revenue = Units sold * price
In this case

$$R(x, y) = (81 - 21x + 17y) * x +  (40 + 11x - 23y) * y$$
$$R(x, y) = 81x - 21x^2 + 17xy +  40y + 11xy - 23y^2 y$$
$$R(x, y) = 81x+40y+28xy-21x^2-23y^2$$ 


* **Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?**

$$R(2.3, 4.1)=81\times 2.3+40\times 4.1+28\times 2.3\times 4.1-21\times (2.3)^2-23\times (4.1)^2 = 116.62$$


#### 4. A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by C(x, y) = $\frac{1}{6} x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?


Consider  x + y = 96   where x  is the number of units produced in Los Angeles and y is the number of units produced in Denver.

Replacing x = 96 - y

$$C(x, y) = C(96 - y, y) = \frac{1}{6} (96 - y)^2 + \frac{1}{6}y^2 + 7(96 - y) + 25y + 700$$

$$=\frac{1}{6} (96 - y)^2 + \frac{1}{6}y^2 + 7(96 - y) + 25y + 700$$

$$=\frac{1}{6} (y^2-192y+9216) + \frac{1}{6}y^2 + (672-7y) + 25y + 700$$

$$=\frac{1}{3} y^2 -14y + 2908$$

$$=C(y)$$

To find the minumum value, $C^{'}(y) = 0$

$$\frac {2}{3} y - 14 = 0$$

$$y = 21$$

Now replacing y value in x = 96 - y = 96 -21


$$x = 75$$

So to minimize the total weekly cost, the number of units produced in Los Angeles should be 75 and in Denver be 21.


#### 5. Evaluate the double integral on the given region

$$\iint_R (e^{8x+3y})dA;R: 2 \leq x \leq and 2 \geq x \geq 4$$


**Write your answer in exact form without decimals.**


$$
\begin{split}
\int_2^4\int_2^4 (e^{8x+3y})\ dy\ dx &= \int_2^4 (\frac{1}{3}e^{8x+3y})|_2^4\ dx\\
&= \int_2^4 ((\frac{1}{3}e^{8x+3*4})-(\frac{1}{3}e^{8x+3*2}))\ dx\\
&= \int_2^4 ((\frac{1}{3}e^{8x+12})-(\frac{1}{3}e^{8x+6}))\ dx\\
&= \int_2^4 \frac{1}{3}e^{8x+6}(e^6-1)\ dx\\
&=  \frac{1}{3*8}e^{8x+6}(e^6-1) |_2^4\\
&= \frac{1}{24}e^{8*4+6}(e^6-1)-\frac{1}{24}e^{8*2+6}(e^6-1)\\
&= \frac{1}{24}(e^6-1)(e^{38}-e^{22})\\
&= \frac{1}{24}(e^{44} - e^{38} - e^{28} + e^{22})
\end{split}
$$















