---
title: "Data605_Assignment7"
author: "Amit Kapoor"
date: "3/11/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.

Given Y denote the minimum of the Xi’s. 

P(Y) = min(X1, X2, X3, ..... Xn)

Assuming each Xi has k possibilities: 1,2,3,....k, then total number of assignments would be $k^{n}$. 
The number of ways of getting X = 1 is k^n - (k - 1)^n / k^n, since k^n is the total number of possibilities and (k-1)^n is all of the options where none of the Xi’s are equal to 1.

P(X=1) = $\frac{k^{n} - (k-1)^{n}}{k^{n}}$

Generalizing
P(X=2) = $\frac{(k-2+1)^{n} - (k-2)^{n}}{k^{n}}$

P(X=3) = $\frac{(k-3+1)^{n} - (k-3)^{n}}{k^{n}}$

.....

P(X=t) = $\frac{(k-t+1)^{n} - (k-t)^{n}}{k^{n}}$



## 2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).


### a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

Geometric Distribution
P(T=n) = $q^{n-1}*p$



```{r prob-geometric, eval=TRUE}
# probability of one failure every ten years
p_fail <- 1/10

# probability of not being failed
p_no_fail <- 1-p_fail

# 8 years
n <- 8

# probability that the machine will fail after 8 years
p_geom <- 1 - pgeom(n-1, p_fail)
p_geom
```


```{r geom-exp, echo=TRUE}

#Expected Value of being failed
exp_val_geom <- 1/p_fail

exp_val_geom
```


```{r geom-sd, echo=TRUE}
#Std deviation
sd_geom <- sqrt(p_no_fail/(p_fail^2))

sd_geom
```



### b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

Exponential Distribution
P(X $\leqslant$ x) = 1 – $e^{–mx}$ where m is decay parameter. 

```{r prob-expo, eval=TRUE}

lambda <- 1/10

# probability that the machine will fail after 8 years
p_expo <- 1 - pexp(n, lambda)
p_expo
```


```{r expo-exp, echo=TRUE}

#Expected Value of being failed
exp_val_expo <- 1/lambda

exp_val_expo
```


```{r expo-sd, echo=TRUE}
#Std deviation
sd_expo <- sqrt(1/(lambda^2))

sd_expo
```

### c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

Binomial Distribution
b(n, p, k) = ${n \choose k}$ * $p^{k}$ * $q^{n-k}$ 
where q = 1-p

```{r prob-binomial,, eval=TRUE}
n <- 8
p <- 1/10
q <- 1-p
k <- 0 # 0 success in 8 years

# probability that the machine will fail after 8 years
p_bino <- dbinom(k, n, p)
p_bino
```


```{r bino-exp, echo=TRUE}

#Expected Value of being failed
exp_val_bino <- n*p

exp_val_bino
```


```{r bino-sd, echo=TRUE}
#Std deviation
sd_bino <- sqrt(n*p*q)

sd_bino
```

### d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.

Poisson Dstribution
P(X=k) = $e^{- \lambda }$ * $\frac{\lambda ^{k}}{k!}$


```{r prob-poisson, eval=TRUE}
lambda <- 8/10
k <- 0

# probability that the machine will fail after 8 years
p_pois <- ppois(k, lambda=lambda)
p_pois
```


```{r pois-exp, echo=TRUE}

#Expected Value of being failed
exp_val_pois <- lambda

exp_val_pois
```


```{r pois-sd, echo=TRUE}
#Std deviation
sd_pois <- sqrt(lambda)

sd_pois
```

















