---
title: "Data605 - Assignment9"
author: "Amit Kapoor"
date: "3/25/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Page 363 #11 The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the nth day of the year. Finn observes that the differences $X_n$ = $Y_{n+1}$ - $Y_n$ appear to be independent random variables with a common distribution having mean $\mu$ = 0 and variance $\sigma^2$ = 1/4. If $Y_1$ = 100, estimate the probability that $Y_{365}$ is

#### (a) $\geq$ 100.

```{r a, echo=TRUE}
mean <- 0
var <- 1/4
pnorm((100-100), mean = mean, sd = sqrt(var*364), lower.tail = FALSE)
```


#### (b) $\geq$ 110.

```{r b, echo=TRUE}

pnorm((110-100), mean = mean, sd = sqrt(var*364), lower.tail = FALSE)
```

#### (c) $\geq$ 120.

```{r c, echo=TRUE}

pnorm((120-100), mean = mean, sd = sqrt(var*364), lower.tail = FALSE)
```




### 2. Calculate the expected value and variance of the binomial distribution using the moment generating function.

Moment Generating Functions is literally the function that generates the moments i.e. E(X), E($X^2$)..
In this case we will find out the first and second derivative to get the expected value and variance.


#### Binomial distribution



$$g(t)=\sum _{ j=0 }^{ n }{ { e }^{ tj } } (n!/(n-j)!j!){ p }^{ j }{ q }^{ n-j }$$
$$g(t)=\sum _{ j=0 }^{ n }{ (n!/(n-j)!j!) } { { (pe }^{ t }) }^{ j }{ q }^{ n-j }$$
$$g(t)={ ({ pe }^{ t }+q) }^{ n }$$

First Derivative:

$$g'(t)=n{ ({ pe }^{ t }+q) }^{ n-1 }{ pe }^{ t }$$

$$g'(0)={ n(p+q) }^{ n-1 }p=np$$

$$\mu ={ \mu  }_{ 1 }=g'(0)=np$$



Second Derivative:

$$g''(t)=n(n-1)({ pe }^{ t }+q){ ({ pe }^{ t }) }^{ 2 }+n{ ({ pe }^{ t }+q) }^{ n }{ pe }^{ t }  $$

$$g''(0)=\quad { n(n-1)(p+q) }^{ n-1 }{ p }^{ 2 }+{ n(p+q) }^{ n }p$$

$$g''(0)=n(n-1){ p }^{ 2 }+np$$



$${ \sigma  }^{ 2 }={ \mu  }_{ 2 }-{ \mu  }_{ 1 }^{ 2 }1=g''(0)-{ g'(0) }^{ 2 }\\ { \sigma  }^{ 2 }=n(n-1){ p }^{ 2 }+np-{ (np) }^{ 2 }\\ { \sigma  }^{ 2 }=np[(n-1)p+1-np]\\ { \sigma  }^{ 2 }=np[(np-p)+1-np]$$

$${ \sigma  }^{ 2 }=np[1-p]$$

$${ \sigma  }^{ 2 }=npq$$

### 3. Calculate the expected value and variance of the exponential distribution using the moment generating function.


#### Exponential distribution


The exponential distribution probability density function is:

$$\lambda { e }^{ -\lambda x }$$
Moment generating function for the binomial distribution is:

$$g(t)=\frac { \lambda }{ \lambda-t } \quad for\quad t<\lambda$$

First Derivative,

$$g'(t)=\frac {\lambda }{ { (\lambda-t) }^{ 2 } } $$
To get mean by evaluating for t=0,

$$g'(0)=\frac { 1 }{\lambda} $$


Second Derivative,

$$g''(t)=\frac { 2 \lambda}{ (\lambda-t)^{ 3 } } $$
To get variance, by evaluating for t=0,

$$g''(0)=\frac {2}{\lambda^2} $$

variance is

$${ \mu  }^{ 2 }-{ \mu  }_{ 1 }^{ 2 }=\frac { 2 }{ { \lambda  }^{ 2 } } -{ (\frac { 1 }{ \lambda  } ) }^{ 2 }=\frac { 2 }{ \lambda ^{ 2 } } -\frac { 1 }{ \lambda ^{ 2 } } =\frac { 1 }{ \lambda ^{ 2 } }  $$



#### Reference

https://towardsdatascience.com/moment-generating-function-explained-27821a739035





