---
title: "Data605 - Discussion10"
author: "Amit Kapoor"
date: "3/31/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Page 414 Ex 7.

Find the matrices $P^2$, $P^3$, $P^4$, and $P^n$ for the Markov chain determined by the transition matrix P = $A = \left( \begin{matrix} 1&0 \\ 0&1 \end{matrix} \right)$. Do the same for the transition matrix P = $A = \left( \begin{matrix} 0&1 \\ 1&0 \end{matrix} \right)$. Interpret what happens in each of these processes.

```{r expm, echo=TRUE}

# Computation of the matrix exponential
library(expm)
```



```{r}

v1 <- c(1,0,0,1)
P1 <- matrix(v1, nrow = 2)
P1
```


```{r}
P1%^%2
```


```{r}
P1%^%3
```




```{r}
P1%^%4
```

Hence if P = $\left( \begin{matrix} 1&0 \\ 0&1 \end{matrix} \right)$ then $P^n$ = P = $\left( \begin{matrix} 1&0 \\ 0&1 \end{matrix} \right)$  as P is an identity matrix. It would be true for any size identity matrix.




```{r}

v2 <- c(0,1,1,0)
P2 <- matrix(v2, nrow = 2)
P2
```


```{r}
P2%^%2
```

```{r}
P2%^%3
```



```{r}
P2%^%4
```



```{r}
P2%^%5
```



Hence if P = $\left( \begin{matrix} 0&1 \\ 1&0 \end{matrix} \right)$ then 

if n is even, $P^n$ = $\left( \begin{matrix} 1&0 \\ 0&1 \end{matrix} \right)$  

if n is odd, $P^n$ = $\left( \begin{matrix} 0&1 \\ 1&0 \end{matrix} \right)$  







