---
title: "Data605_Assignment5"
author: "Amit Kapoor"
date: "2/26/2020"
output:
  pdf_document: default
  html_document:
    toc: true
    toc_depth: 3
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square.

```{r B-C, echo=TRUE}
set.seed(67)
n <- 10000
B <- runif(n, min = 0, max = 1)
C <- runif(n, min = 0, max = 1)
```


```{r min-max-B, echo=TRUE}
cat("min(B)->", format(min(B), scientific = F), "max(B)->",max(B))
```


```{r min-max-C, echo=TRUE}
cat("min(c)->", format(min(C), scientific = F), "max(C)->",max(C))
```


```{r hist-B, echo=TRUE}

# skewness of B
hist(B, probability = T, main = 'B histogram')
```

```{r hist-C, echo=TRUE}

# skewness of C
hist(C, probability = T, main = 'C histogram')
```

Since the bins of both the histograms are evenly distributed, we can conclude B & C are proper probability distributions.




## Find the probability that 

### B+C < 1/2



```{r prob-1, echo=TRUE}

# The probability of B+C less than 1/2 is
a <- sum((B+C) < 0.5) / n
print(paste("The probability of B+C less than 1/2 is",a))
```



### BC < 1/2

```{r prob-2, echo=TRUE}

# The probability of B*C less than 1/2 is
b <- sum((B*C) < 0.5) / n
print(paste("The probability of BC less than 1/2 is",b))
```


### |B-C| < 1/2

```{r prob-3, echo=TRUE}

# The probability of |B-C| less than 1/2 is
c <- sum(abs(B-C) < 0.5) / n
print(paste("The probability of |B,C| less than 1/2 is",c))
```



### max{B,C} < 1/2


```{r prob-4, echo=TRUE}

# The probability of max{B,C} less than 1/2 is
e <- sum(pmax(B,C) < 0.5) / n
print(paste("The probability of max{B,C} less than 1/2 is",e))
```



### min{B,C} < 1/2

```{r prob-5, echo=TRUE}

# The probability of min{B,C} less than 1/2 is
f <- sum(pmin(B,C) < 0.5) / n
print(paste("The probability of min{B,C} less than 1/2 is",f))
```




