---
title: "Data605 - Assignment10"
author: "Amit Kapoor"
date: "4/1/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if

### (a) he bets 1 dollar each time (timid strategy).

$$P = \frac { 1-{ (\frac { q }{ p } ) }^{ s } }{ 1-{ (\frac { q }{ p } ) }^{ M } }  $$

$$P = \frac { 1-{ (\frac { 0.6 }{ 0.4 } ) }^{ 1 } }{ 1-{ (\frac { 0.6 }{ 0.4 } ) }^{ 8 } }  $$

```{r}
p <- 0.4
q <- 0.6
r<- q/p
s <- 1
M <- 8

for(n in s:M) {
  P <- round( (1-r^n)/(1-r^M), 4)
  print(P)
} 
```

hence $P_{timid}$ = 0.0203


### (b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).

Here Smith bets the entire amount (up to 8). If he wins, he will follow the following sequence: 1,2,4,8.

He starts with 1 dollar and must win 3 times in a row with probability 0.4. So, the probability of 3 successes in a row is

$P_{bold}$ = $(0.4)^3$ = 0.064


### (c) Which strategy gives Smith the better chance of getting out of jail?

The bold strategy seems better as compared to timid strategy for Smith to come out of jail.



