---
title: "Data 605 - Assginment 2"
author: "Amit Kapoor"
date: "2/8/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Problem set 1

## (1) Show that $A$ $\mathbf{A}^\intercal$ $\neq$ $\mathbf{A}^\intercal$ $A$ in general. (Proof and demonstration.)

### Proof:-

Let A be a m x n matrix. By definition its transpose $\mathbf{A}^\intercal$ will be n x m matrix.

Given the definition of matrix multiplication, multiplying A to $\mathbf{A}^\intercal$ will return a m x m square matrix.

Similarly multiplying $\mathbf{A}^\intercal$ to A will return a n x n matix.

Therefore if m $\neq$ n then result of $A$ $\mathbf{A}^\intercal$ dimension would be different of $\mathbf{A}^\intercal$ $A$

hence $A$ $\mathbf{A}^\intercal$ $\neq$ $\mathbf{A}^\intercal$ $A$


### Demonstration:-

$A = \left( \begin{matrix} 1&2&3\\ 4&5&6 \end{matrix} \right)$
$\mathbf{A}^\intercal$ = $\left( \begin{matrix} 1&4\\ 2&5\\ 3&6 \end{matrix} \right)$

$A$ $\mathbf{A}^\intercal$ = $\left( \begin{matrix} 1&2&3\\ 4&5&6 \end{matrix} \right)$ x $\left( \begin{matrix} 1&4\\ 2&5\\ 3&6 \end{matrix} \right)$ = $\left( \begin{matrix} 14&32\\ 32&77 \end{matrix} \right)$

$\mathbf{A}^\intercal$ $A$ = $\left( \begin{matrix} 1&4\\ 2&5\\ 3&6 \end{matrix} \right)$ x $\left( \begin{matrix} 1&2&3\\ 4&5&6 \end{matrix} \right)$ = $\left( \begin{matrix} 17&22&27\\ 22&29&36\\ 27&36&45 \end{matrix} \right)$

hence $A$ $\mathbf{A}^\intercal$ $\neq$ $\mathbf{A}^\intercal$ $A$

## (2) For a special type of square matrix A, we get $A$ $\mathbf{A}^\intercal$ $\neq$ $\mathbf{A}^\intercal$ $A$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

For symmetric matrix $A$ = $\mathbf{A}^\intercal$ then $A$ $\mathbf{A}^\intercal$ = $\mathbf{A}^\intercal$ $A$

matrix A is symmetric if $a_{ij}$ = $a_{ji}$ for all i,j

### Example:-

A = $\left( \begin{matrix} 1&2&3\\ 2&3&4 \\ 3&4&7 \end{matrix} \right)$
$\mathbf{A}^\intercal$ = $\left( \begin{matrix} 1&2&3\\ 2&3&4 \\ 3&4&7 \end{matrix} \right)$

$A$ $\mathbf{A}^\intercal$ = $\mathbf{A}^\intercal$ $A$ = $\left( \begin{matrix} 1&2&3\\ 2&3&4 \\ 3&4&7 \end{matrix} \right)$ x $\left( \begin{matrix} 1&2&3\\ 2&3&4 \\ 3&4&7 \end{matrix} \right)$ = $\left( \begin{matrix} 14&20&32\\ 20&29&46 \\ 3&46&74 \end{matrix} \right)$


# 1. Problem set 2

## Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.

## Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

```{r lu-function, echo=TRUE}

library(matrixcalc)

LU_factorization <- function(A) {
  
  # To check first if given matrix is a square matrix
  if(is.square.matrix(A) == FALSE) {
    return("Only square matrix is allowed here.")
  }
  
  U <- A
  n <- dim(A)[1]
  # identity matrix
  L <- diag(n)  
  
  if(n==1){
    return(list(L,U))
  }
  
  for(i in 2:n){
    for(j in 1:(i-1)) {
      m <- -U[i,j] / U[j,j]
      U[i,] <- m * U[j,] + U[i,]
      L[i,j] <- -m
    }
  }
  return(list(L,U))
}
```




```{r matrix-to-test, echo=TRUE}
A <- matrix(c(1,2,3,
              4,5,6,
              7,8,9), byrow = T, ncol = 3)
A
```

```{r test-lu-function, echo=TRUE}
lu <- LU_factorization(A)
lu
```



